Parsing Postfix log files

### 2008-07-23 doc/logparser-conclusion.tex
Improved parts of the efficiency section and inlined graphs. Fixed up some references; added $Id$ to front page.

Given the diverse nature of the log lines and the ease with which
administrators can cause new lines to be logged (\sectionref{postfix
background}), enabling users to easily extend the parser to deal with new
log lines is a design imperative (\sectionref{parser design}).  Despite the
resulting initial increase in complexity the task is quite tractable,
though it does raise interesting efficiency and optimisation questions
(answered in \sectionref{rule efficiency}).  Providing a tool
(\sectionref{creating new rules}) to ease the generation of \regexes{} from
unparsed log lines should greatly help users add rules to parse formerly
unparsed log lines.

>>>>

Given the diverse nature of the log lines and the ease with which
administrators can cause new lines to be logged (\sectionref{postfix
background}), enabling users to easily extend the parser to deal with new
log lines is a design imperative (\sectionref{parser design}).  Despite the
resulting initial increase in complexity the task is quite tractable,
though it does raise interesting efficiency and optimisation questions
(answered in \sectionref{parser efficiency}).  Providing a tool
(\sectionref{creating new rules}) to ease the generation of \regexes{} from
unparsed log lines should greatly help users add rules to parse formerly
unparsed log lines.

### 2008-07-23 doc/logparser-implementation.tex
Improved parts of the efficiency section and inlined graphs. Fixed up some references; added $Id$ to front page.

        For efficiency the keywords are expanded and every rule's \regex{}
        is compiled before attempting to parse the log file --- otherwise
        each \regex{} would be recompiled each time it was used, resulting
        in a large, data dependent slowdown.  Rule efficiency concerns are
        discussed in \sectionref{rule efficiency}, with the impact of
        caching compiled \regexes{} covered in \sectionref{Caching each
        regex}.

>>>>

        For efficiency the keywords are expanded and every rule's \regex{}
        is compiled before attempting to parse the log file --- otherwise
        each \regex{} would be recompiled each time it was used, resulting
        in a large, data dependent slowdown.  Efficiency concerns are
        discussed in \sectionref{parser efficiency}, with the impact of
        caching compiled \regexes{} covered in \sectionref{Caching each
        regex}.

### 2008-07-23 doc/logparser-results.tex
Improved parts of the efficiency section and inlined graphs. Fixed up some references; added $Id$ to front page.

\subsection{Efficiency}

\subsubsection{Rule efficiency}

\label{rule efficiency}

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
deal with 75 MB log files containing 300,000 log lines (generated daily on
a mail server handling mail for approximately 700 users --- large scale
mail servers would have much larger log files on a daily basis).  When
generating the data for the graphs included in \sectionref{graphs},
\numberOFlogFILES{} log files (totaling 10.08 GB, \numberOFlogLINEShuman{}
log lines) were each parsed 10 times, the first run discarded, and the
execution time for the remaining 9 runs averaged.  The first run is
discarded for two reasons:

@@

An important property of a parser is how execution time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{execution time vs file size vs number of lines graph}
shows the execution time in seconds, file size in MB and tens of thousands
of log lines per log file.  All three lines run roughly in parallel, giving
a visual impression that the algorithm scales linearly with input size.
This impression is borne out by \graphref{execution time vs file
size vs number lines factor} which plots the ratio of file size vs
execution time and ratio of number of log lines vs execution time (higher
is better); \tableref{execution time vs file size vs number lines
factor table} shows the ratios for different breakdowns of the log files.
As the reader can see the ratios are quite tightly banded, showing that the
algorithm scales linearly: the much larger log files between points 60 and
70 on the X axis in \graphref{execution time vs file size vs
number of lines graph} actually cause the ratio to increase (i.e.\
improve), rather than decrease.  The strange behaviour where larger log
files are parsed more efficiently is explained fully in \sectionref{Why are
there dips in the graphs?}.

@@

Rule ordering was mentioned in \sectionref{rule attributes} and will be
covered in greater detail in this section.  At the time of writing there
are \numberOFrules{} different rules: the top 10\% match the vast majority
of log lines, with the remaining log lines split across the other 90\% of
the rules (as shown in \graphref{rule hits graph}).  Assuming that
the distribution of log lines is reasonably steady over time, program
efficiency should benefit from trying more frequently matching rules before
those which match less frequently.  To test this hypothesis three full test
runs were performed with different rule orderings:

@@

Graphs~\refwithpage{percentage increase of shuffled over normal}
and~\refwithpage{percentage increase of reversed over normal} show the
percentage increase of execution times, with \tableref{Execution
time increase for different rule orderings} showing the mean increases for
different groupings of log files.  Overall this provides a modest but
worthwhile performance increase of approximately 9\%, for a small
investment in time and programming.

@@

\renewcommand{\figurename}{Graph}

\subsubsection{Introduction}

Graphs are an excellent means of displaying data, transforming a
meaningless stream of numbers into an easily comprehensible form, where
anomalies and patterns are immediately obvious.  These graphs are used to
illustrate the topics discussed in \sectionref{rule efficiency}.  The
graphs in the first section cover parser scalability, demonstrating that
performance scales linearly with input size.  The impact of rule ordering
is shown in \sectionref{rule ordering graphs}, and the anomalous dips and
peaks apparent in some graphs are explained.  The third group of graphs
vividly shows the huge impact that caching compiled \regexes{} has on
parser performance.  Miscellaneous graphs are presented in the final group
of graphs; these graphs are referenced from various places in the test.
This section concludes with a breakdown of the rule hits accumulated during
a single test run of the parser.

\newpage

\subsubsection{Parser scalability}

\Graphref{execution time vs file size vs number lines factor}
shows the ratio of execution time vs file size and number of lines (higher
is better, it means more bytes or lines processed per second).  The ratios
are quite tightly banded with the exception of log files 22 and 62--68,
where they are noticeably higher; \graphref{execution time vs file
size vs number lines factor} and \tableref{execution time vs file
size vs number lines factor table} show that the parser's execution time
scales linearly with input size.

\subsubsection{Rule ordering}

@@

The graphs presented in this section illustrate the topics discussed in
\sectionref{rule efficiency}.  The first collection of graphs are about
parser scalability, showing the linear relationship between execution time
and input size.  \sectionref{rule ordering graphs} demonstrates the effect
of rule ordering on execution time, and the unexpected consequences of
specific inputs.  The necessity of caching compiled \regexes{} is attested
to by the third group of graphs, where the difference between caching and
not caching is staggering.  The penultimate section contains a breakdown of
the rule hits accumulated during a single test run of the parser.
Miscellaneous graphs expected to be useful are collected in the final
section.

>>>>

\subsection{Parser Efficiency}

\label{parser efficiency}

@@

Parsing efficiency is an obvious concern when the parser routinely needs to
deal with 75 MB log files containing 300,000 log lines (generated daily on
a mail server handling mail for approximately 700 users --- large scale
mail servers would have much larger log files on a daily basis).  When
generating the timing data used in this section, \numberOFlogFILES{} log
files (totaling 10.08 GB, \numberOFlogLINEShuman{} log lines) were each
parsed 10 times, the results of the first first run discarded, and the
execution time for the remaining 9 runs averaged.  The first run is
discarded for two reasons:

@@

An important property of a parser is how execution time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{execution time vs file size vs number of lines graph} shows the
execution time in seconds, file size in MB, and tens of thousands of log
lines per log file.  All three lines run roughly in parallel, giving the
impression that the algorithm scales linearly with input size.  This
impression is borne out by \graphref{execution time vs file size vs number
lines factor} which plots both the ratio of file size vs execution time,
and the ratio of number of log lines vs execution time (higher is better in
both cases); \tableref{execution time vs file size vs number lines factor
table} shows the ratios for different groups of log files.  As the
reader can see the ratios are quite tightly banded, showing that the
algorithm scales linearly: the much larger log files between points 60 and
70 on the X axis in \graphref{execution time vs file size vs number of
lines graph} actually cause the ratio to increase (i.e.\ improve), rather
than decrease.  The strange behaviour where larger log files are parsed
more efficiently is explained fully in \sectionref{Why are there dips in
the graphs?}.

@@

At the time of writing there are \numberOFrules{} different rules: the top
10\% match the vast majority of log lines, with the remaining log lines
split across the other 90\% of the rules (as shown in \graphref{rule hits
graph}).  Assuming that the distribution of log lines is reasonably steady
over time, parser efficiency should benefit from trying more frequently
matching rules before those which match less frequently.  To test this
hypothesis three full test runs were performed with different rule
orderings:

@@

\Graphref{percentage increase of shuffled over normal} shows the percentage
increase of shuffled over normal, with the percentage increase of reverse
over normal following in \graphref{percentage increase of reversed over
normal}; the mean increases for different groupings of log files are given
in \tableref{Execution time increase for different rule orderings}.
Overall this optimisation provides a modest but worthwhile performance
increase of approximately 10\%, for a small investment in time and
programming.

XXX SHOW NORMAL AS A REDUCTION FROM SHUFFLE, AND REVERSE AS AN INCREASE
OVER SHUFFLE\@.

@@

Graphs are an excellent means of displaying data, transforming a
meaningless stream of numbers into an easily comprehensible form, where
anomalies and patterns are immediately obvious.  The
graphs in the first section cover parser scalability, demonstrating that
performance scales linearly with input size.  The impact of rule ordering
is shown in \sectionref{rule ordering graphs}, and the anomalous dips and
peaks apparent in some graphs are explained.  The third group of graphs
vividly shows the huge impact that caching compiled \regexes{} has on
parser performance.  Miscellaneous graphs are presented in the final group
of graphs; these graphs are referenced from various places in the test.
This section concludes with a breakdown of the rule hits accumulated during
a single test run of the parser.

@@

The first collection of graphs are about
parser scalability, showing the linear relationship between execution time
and input size.  \sectionref{rule ordering graphs} demonstrates the effect
of rule ordering on execution time, and the unexpected consequences of
specific inputs.  The necessity of caching compiled \regexes{} is attested
to by the third group of graphs, where the difference between caching and
not caching is staggering.  The penultimate section contains a breakdown of
the rule hits accumulated during a single test run of the parser.
Miscellaneous graphs expected to be useful are collected in the final
section.

### 2008-07-23 doc/logparser.tex
More proofreading changes.
Merged 'Other Postfix log parsers reviewed' and 'Previous research in this area' to create a 'State of the Art Review', which is now section 3.
Removed 'Throughput monitor' review.

Improved introduction and summary in 'State of the Art Review'; miscellaneous
related changes.

Moved 'Graphs' section closer to where it will eventually be, the 'Results'
section.
Moved the architecture stuff into it's own section.
Collected all the implementation bits into one section.
Gathered all the results bits into their own section.
Removed empty files.
Improved parts of the effic

% Produce nicer references like "section 4.3 on the next page"
% Must be loaded before hyperref.
\usepackage{varioref}
% Warnings rather than errors when references cross page boundaries.
\vrefwarning{}
% Errors when loops are encountered
\vrefshowerrors{}
% Shorten the text used for references with page numbers.
\renewcommand{\reftextfaraway}[1]{%
    [p.~\pageref{#1}]%
}
% Replacement for \ref{}, adds the page number too.
\newcommand{\refwithpage}[1]{%
    \empty{}\vref{#1}%
    %\empty{}\ref{#1} [p.~\pageref{#1}]%
}
% section references, automatically add \textsection
\newcommand{\sectionref}[1]{%
    \textsection{}\vref*{#1}%
    %\textsection{}\refwithpage{#1}%
}
% A reference with a space between the label and the reference.
\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\tableref}[1]{%
    \refwithlabel{table}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}
% graph references
\newcommand{\graphref}[1]{%
    \refwithlabel{graph}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Graph}{#1}%
}
% figure references
\newcommand{\figureref}[1]{%
    \refwithlabel{figure}{#1}%
}
\newcommand{\Figureref}[1]{%
    \refwithlabel{Figure}{#1}%
}

@@

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[3]{
    \begin{figure}[hbt!]
        \caption{#2}\label{#3}
        \includegraphics{#1}
    \end{figure}
}

%\showtable{filename}{caption}{label}
\newcommand{\showtable}[3]{
    \begin{table}[ht]
        \caption{#2}\label{#3}
        \input{#1}
    \end{table}
}

@@

\title{Parsing Postfix log files}
\author{John Tobin \\ School of Computer Science and Statistics \\
Trinity College \\ Dublin 2 \\ Ireland \\ tobinjt@cs.tcd.ie}
\maketitle

@@

XXX ``.'' AFTER ETC\@; SHOULD I HAVE A BACKSLASH AFTER THE ``.''?

@@

\begin{verbatim}
Introduction
Background
State of the Art Review
Parser Architecture
Postfix Parser Implementation
    Database Schema
    Rules
    Actions
    Complications
Results - BETTER TITLE?
    Coverage
    Efficiency
    Limitations and Improvements
Conclusion
References
Glossary
Acronyms
\end{verbatim}

@@

\newpage
\input{logparser-introduction.tex}
\input{logparser-background.tex}
\input{logparser-database-schema.tex}
\input{logparser-parsing-rules.tex}
\input{logparser-parsing-algorithm.tex}
\input{logparser-coverage.tex}
\input{logparser-limitations.tex}
\input{logparser-conclusion.tex}

@@

\input{logparser-other-parsers-review.tex}

@@

\input{logparser-graphs.tex}

>>>>

% Embed SVN Id etc.
\usepackage{svn}

% Produce nicer references like "section 4.3 on the next page"
% Must be loaded before hyperref.
\usepackage{varioref}
% Warnings rather than errors when references cross page boundaries.
\vrefwarning{}
% Errors when loops are encountered
\vrefshowerrors{}
% Shorten the text used for references with page numbers.
\renewcommand{\reftextfaraway}[1]{%
    [p.~\pageref{#1}]%
}
% Replacement for \ref{}, adds the page number too.
\newcommand{\refwithpage}[1]{%
    \empty{}\vref{#1}%
    %\empty{}\ref{#1} [p.~\pageref{#1}]%
}
% section references, automatically add \textsection
\newcommand{\sectionref}[1]{%
    \textsection{}\vref*{#1}%
    %\textsection{}\refwithpage{#1}%
}
% A reference with a space between the label and the reference.
\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\tableref}[1]{%
    \refwithlabel{table}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}
% graph references
% SHOULD THESE BE LABELLED WITH graph OR figure?
\newcommand{\graphref}[1]{%
    \refwithlabel{graph}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Graph}{#1}%
}
% figure references
\newcommand{\figureref}[1]{%
    \refwithlabel{figure}{#1}%
}
\newcommand{\Figureref}[1]{%
    \refwithlabel{Figure}{#1}%
}

@@

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[3]{%
    \begin{figure}[hbt!]
        \caption{#2}\label{#3}
        \includegraphics{#1}
    \end{figure}
}

%\showtable{filename}{caption}{label}
\newcommand{\showtable}[3]{%
    \begin{table}[ht]
        \caption{#2}\label{#3}
        \input{#1}
    \end{table}
}

\newcommand{\tableline}[0]{%
    \noalign{\smallskip}%
    \hline%
    \noalign{\smallskip}%
}

@@

% WSUIPA fonts.
\input{ipamacs}

\title{Parsing Postfix log files}
\author{John Tobin \\ School of Computer Science and Statistics \\
Trinity College \\ Dublin 2 \\ Ireland \\ tobinjt@cs.tcd.ie}
\maketitle
\SVN$Id$
\SVNId{}

@@

@@

\begin{verbatim}
Introduction
Background
State of the Art Review
Parser Architecture
Postfix Parser Implementation
    Database Schema
    Rules
    Actions
    Complications
    Limitations and Improvements
Results - BETTER TITLE?
    Coverage
    Efficiency
Conclusion
References
Glossary
Acronyms
\end{verbatim}

@@

\newpage
\input{logparser-introduction.tex}
\input{logparser-background.tex}
\input{logparser-other-parsers-review.tex}
\input{logparser-architecture.tex}
\input{logparser-implementation.tex}
\input{logparser-parsing-rules.tex}
\input{logparser-results.tex}
\input{logparser-conclusion.tex}

@@

@@
